---
title: "Sapflow data analysis"
author: "Aji John"
date: "2/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(lubridate)
```

## R Markdown for analyzing the data from the sensors

Put the data files in the folder 'data' in the current directory.

```{r cars}
location <- '.'

# Storing the column names because we might have a case where file headers are not written.

colNamesSapflow <- c("unix_timestamp" ,"human_timestamp"  ,        "nodeID"    ,              
  "temp1"    ,                "temp2"  ,                  "millisSinceHeatPulse"   , 
"outsideTemp"        ,      "outsideHumidity"      ,    "soilMoisture"  ,          
 "rtcUnixTimestamp"    ,     "millisSinceReferenceTemp")

# get list of files 
files <- list.files(paste(location, "data", sep="/"), pattern=".csv", full.names=TRUE)

```

Read and merge all the dataframes (one dataframe per CSV)
```{r cars}
df <- data.frame()

for ( file in files ) {
  file_df <- read_csv(file)
  # Not ideal, but skipping for now, these the files where we don't have the header
  # TODO support parsing the files whicxh don't have the header
  if ((colnames(file_df)[1] == "unix_timestamp"))
    df <- rbind(df,file_df)
}

```

## Parse the dates
Complicated but essential part, human timestamp is bound to be in UTC, so converting to local timse zone (Pacific)

```{r }

# Join them all into a single data frame
sensor_data <- df
sensor_data$human_timestamp<- strptime(sensor_data$human_timestamp, "%m/%d/%Y %H:%M:%S",tz="Europe/London")
sensor_data$human_timestamp_pst<-format(as.POSIXct(sensor_data$human_timestamp),tz="America/Los_Angeles",usetz=TRUE)
sensor_data$human_timestamp_pst<- as.POSIXct(sensor_data$human_timestamp_pst)

```

Storing , month and day.

```{r }


sensor_data$day<- day(sensor_data$human_timestamp_pst)
sensor_data$month<- month(sensor_data$human_timestamp_pst)
```

Plot the data, spot check. Ideally, don't want to see +40 or -40, issue with sensor

```{r }

# Join them all into a single data frame
sensor_data %>% select(human_timestamp_pst,temp1,temp2,day) %>% filter(day ==7)  %>% 
  ggplot() + 
geom_line( aes(x=human_timestamp_pst,y=temp1, color = "TC1")) +
geom_line( aes(x=human_timestamp_pst,y=temp2, color = "TC2"))

```
Another way to validate , we want ratios to be above 0. Sensor malfunction if it is below 0.

```{r }

# Join them all into a single data frame
sensor_data %>% select(human_timestamp_pst,temp1,temp2,day)   %>% 
  ggplot() + 
geom_histogram( aes(temp1/temp2),binwidth = .5) + xlim(-2,2) + 
  facet_grid(.~day)+
  theme_minimal(base_size=24) +
   geom_vline(xintercept = 0, linetype="dotted", 
                color = "red", size=1) +
  labs(x='Ratio of Temp 1 and 2',y='Count',title = "")
 ggsave("Sanity-check.png",dpi=300, dev='png', height=12, width=18, units="in")

```


