---
title: "Sapflow"
author: "Aji John"
date: "12/16/2020"
output: html_document
---

# Running notebook for the Sapflow data captured by Sapflow nodes (DotmoteLabs)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Sapflow measurements help with tree water use that can be exteded to canopy level transpiration assessments.

```{r }
library(knitr)
library(tidyverse)
library(lubridate)
```

## Load helper functions


```{r setup, include=FALSE}
## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```

## Verify the input files


```{r pressure, echo=FALSE}
sapflow_sh_2020_2<-  read_csv('./data/sapflow-dotmote-labs-2_2020-10-07_to_2020-12-07.csv')

# Tell R studio that timestamp is already in PST
sapflow_sh_2020_2$dateP <- as_datetime(sapflow_sh_2020_2$timestamp, tz="America/Los_Angeles")

sapflow_sh_2020_3<-  read_csv('./data/sapflow-dotmote-labs-3_2020-10-07_to_2020-12-07.csv')

# Tell R studio that timestamp is already in PST
sapflow_sh_2020_3$dateP <- as_datetime(sapflow_sh_2020_3$timestamp, tz="America/Los_Angeles")

sapflow_sh_2020_4<-  read_csv('./data/sapflow-dotmote-labs-4_2020-10-07_to_2020-12-07.csv')

# Tell R studio that timestamp is already in PST
sapflow_sh_2020_4$dateP <- as_datetime(sapflow_sh_2020_4$timestamp, tz="America/Los_Angeles")

sapflow_sh_2020_5<-  read_csv('./data/sapflow-dotmote-labs-5_2020-10-07_to_2020-12-07.csv')

# Tell R studio that timestamp is already in PST
sapflow_sh_2020_5$dateP <- as_datetime(sapflow_sh_2020_5$timestamp, tz="America/Los_Angeles")


```

## Run summary analysis

```{r }
sapflow_sh_2020 <-  rbind(sapflow_sh_2020_2,
                          sapflow_sh_2020_3,
                          sapflow_sh_2020_4,
                          sapflow_sh_2020_5)

  
```

## Plot the raw heat ratio data 

```{r }
sapflow_sh_2020 %>% mutate(meanhr = meanHeatRatio) %>%
  filter(!meanhr =='NaN') %>% mutate(mr=as.numeric(meanhr)) %>%
     mutate(node = as.factor(nodeId),Vh = log(mr),hr = hour(date)) %>%
      filter(date(dateP) == "2020-12-01" ) %>%
  #filter(nodeId %in% c("sapflow-dotmote-labs-2" ,"sapflow-dotmote-labs-3", "sapflow-dotmote-labs-4","sapflow-dotmote-labs-5")) %>%
  ggplot() + geom_point(aes(dateP,mr,color=node) )  +
  facet_grid(~node) +
  theme_classic() + labs(x="Date",y="Mean HR", title = "Node 2-5 SUH - WA ", caption="Source : dotmotelabs.com") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

## Verify diurnal 


## Do zero offsetting 

Plot alternatively, but do offsetting by zeroing out the heatratio by using the values from hrs 0,1 and 2. Assumption being around hours (0-2), ideally no sapflow in plants.

```{r , echo=FALSE}

suh_2_offset <- sapflow_sh_2020 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-dotmote-labs-2"))  %>% 
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))


suh_3_offset <- sapflow_sh_2020 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-dotmote-labs-3"))  %>%
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))

suh_4_offset <- sapflow_sh_2020 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-dotmote-labs-4"))  %>% 
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))


suh_5_offset <- sapflow_sh_2020 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-dotmote-labs-5"))  %>%
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))



sapflow_sh_2020_adjusted <- sapflow_sh_2020
sapflow_sh_2020_adjusted$hr = hour(sapflow_sh_2020$dateP)
sapflow_sh_2020_adjusted$mr <- 0

sapflow_sh_2020_adjusted[sapflow_sh_2020_adjusted$nodeId == "sapflow-dotmote-labs-2",]$mr =  sapflow_sh_2020_adjusted[sapflow_sh_2020$nodeId == "sapflow-dotmote-labs-2",]$meanHeatRatio + (1-as.numeric(suh_2_offset))

sapflow_sh_2020_adjusted[sapflow_sh_2020_adjusted$nodeId == "sapflow-dotmote-labs-3",]$mr =  sapflow_sh_2020_adjusted[sapflow_sh_2020$nodeId == "sapflow-dotmote-labs-3",]$meanHeatRatio + (1-as.numeric(suh_3_offset))

sapflow_sh_2020_adjusted[sapflow_sh_2020_adjusted$nodeId == "sapflow-dotmote-labs-4",]$mr =  sapflow_sh_2020_adjusted[sapflow_sh_2020$nodeId == "sapflow-dotmote-labs-4",]$meanHeatRatio + (1-as.numeric(suh_4_offset))

sapflow_sh_2020_adjusted[sapflow_sh_2020_adjusted$nodeId == "sapflow-dotmote-labs-5",]$mr =  sapflow_sh_2020_adjusted[sapflow_sh_2020$nodeId == "sapflow-dotmote-labs-5",]$meanHeatRatio + (1-as.numeric(suh_5_offset))




#SE 
sapflow_sh_2020_adjusted_se <- summarySE(sapflow_sh_2020_adjusted, measurevar="mr", groupvars=c("nodeId","hr"))




# TODO VPD SE


```

## Plot the adjusted with original 

```{r , echo=FALSE}

sapflow_sh_2020 %>% 
  mutate(node = as.factor(nodeId),hr = hour(dateP)) %>%
  mutate(mr = meanHeatRatio) %>%
  ggplot() + geom_point(aes(hr,mr,color=node) )  +
  facet_grid(~nodeId) +
  theme_classic() + labs(x="Date",y="Mean HR", title = "Node 2-5 SUH - WA  ", caption="Source : dotmotelabs.com") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

```

```{r , echo=FALSE}

sapflow_sh_2020_adjusted %>% 
  mutate(node = as.factor(nodeId),hr = hour(dateP)) %>%
  ggplot() + geom_point(aes(hr,mr,color=node) )  +
  facet_grid(~nodeId) +
  theme_classic() + labs(x="Date",y="Mean HR", title = "Node 2-5 SUH - WA  ", caption="Source : dotmotelabs.com") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

```

## Compare Original and Adjusted HR


```{r , echo=FALSE}

lm(sapflow_sh_2020_adjusted$meanHeatRatio~sapflow_sh_2020_adjusted$mr)

sapflow_sh_2020_adjusted %>% 
  ggplot() +
  geom_point(aes(meanHeatRatio,mr,color=nodeId)) +
  theme_minimal(base_size = 14)
```

## Heat pulse velocity

```{r , echo=FALSE}

sapflow_sh_2020_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
#        filter(date(dateP) == "2020-12-01" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=sv,color=nodeId)) +
   geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = c(.17, 1))+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))

```


## Mass flow


```{r , echo=FALSE}

sapflow_sh_2020_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
   mutate (vs =1.7 * sv )%>% 
   mutate (flow = 3.14 * (1.5)^2 *vs )%>% 
        filter(date(dateP) == "2020-12-01" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=flow,color=nodeId)) +
   geom_line(aes(x = dateP, y=flow,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    #name = expression(~V[h] (cm~h^{"-1"}))
       # Features of the first axis
    name = expression("Sapflow rate"~J~(g~h^{"-1"})),
  ) +
  theme(legend.position = c(.17, 1))+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))

```

## Map the patterns for Sunny days with Cloudy/Rainy Days - Use Agrimet. Use Sapflow velocity curves. Maybe above a threshold of 10 are viable days. 
