---
title: "Sapflow - USDA Gravimetric"
author: "Aji John@DotMote Labs"
date: "05/01/2021"
output:
  pdf_document: default
  html_document: default
---

# Running notebook for the Sapflow data captured by Sapflow nodes (DotmoteLabs)

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Sapflow measurements help with tree water use that can be exteded to canopy level transpiration assessments.

```{r }
library(knitr)
library(tidyverse)
library(lubridate)
```

## Load helper functions


```{r, include=FALSE}
## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```

## Verify the input files


```{r pressure, echo=FALSE}
sapflow_usda_2021_7<-  read_csv('./data/sapflow-usda-7_2021-04-16_to_2021-05-01.csv')

# Tell R studio that timestamp is already in MDT
sapflow_usda_2021_7$dateP <- as_datetime(sapflow_usda_2021_7$timestamp, tz="America/Denver")

sapflow_usda_2021_8<-  read_csv('./data/sapflow-usda-8_2021-04-16_to_2021-05-01.csv')

# Tell R studio that timestamp is already in MDT
sapflow_usda_2021_8$dateP <- as_datetime(sapflow_usda_2021_8$timestamp, tz="America/Denver")

sapflow_usda_2021_9<-  read_csv('./data/sapflow-usda-9_2021-04-16_to_2021-05-01.csv')

# Tell R studio that timestamp is already in MDT
sapflow_usda_2021_9$dateP <- as_datetime(sapflow_usda_2021_9$timestamp, tz="America/Denver")

sapflow_usda_2021_10<-  read_csv('./data/sapflow-usda-10_2021-04-16_to_2021-05-01.csv')

# Tell R studio that timestamp is already in MDT
sapflow_usda_2021_10$dateP <- as_datetime(sapflow_usda_2021_10$timestamp, tz="America/Denver")



```

## Run summary analysis

```{r }
sapflow_usda_2021 <-  rbind(sapflow_usda_2021_7,
                            sapflow_usda_2021_8,
                            sapflow_usda_2021_9,
                            sapflow_usda_2021_10)

  
```

## Plot the raw heat ratio data 

```{r }
sapflow_usda_2021 %>% mutate(meanhr = meanHeatRatio) %>%
  filter(!meanhr =='NaN') %>% mutate(mr=as.numeric(meanhr)) %>%
     mutate(node = as.factor(nodeId),Vh = log(mr),hr = hour(date)) %>%
      filter(date(dateP) == "2021-04-25" ) %>%
  #filter(nodeId %in% c("sapflow-dotmote-labs-2" ,"sapflow-dotmote-labs-3", "sapflow-dotmote-labs-4","sapflow-dotmote-labs-5")) %>%
  ggplot() + geom_point(aes(dateP,mr,color=node) )  +
  facet_grid(~node) +
  theme_classic() + labs(x="Date",y="Mean HR", title = "Node 7-10 USDA - Gravimetric Maize", caption="Source : dotmotelabs.com") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

## Verify diurnal 


## Do zero offsetting 

Plot alternatively, but do offsetting by zeroing out the heatratio by using the values from hrs 0,1 and 2. Assumption being around hours (0-2), ideally no sapflow in plants.

```{r , echo=TRUE}

usda_7_offset <- sapflow_usda_2021 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-usda-7"))  %>% 
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))


usda_8_offset <- sapflow_usda_2021 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-usda-8"))  %>%
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))

usda_9_offset <- sapflow_usda_2021 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-usda-9"))  %>% 
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))

usda_10_offset <- sapflow_usda_2021 %>%  mutate(meanhr = meanHeatRatio,hr = hour(dateP)) %>%
  filter(!meanhr =='NaN') %>% 
  filter(nodeId %in% c("sapflow-usda-10"))  %>% 
  filter( hr %in% c(0,1,2)) %>% summarise(mhr = mean(meanhr))


sapflow_usda_2021_adjusted <- sapflow_usda_2021
sapflow_usda_2021_adjusted$hr = hour(sapflow_usda_2021$dateP)
sapflow_usda_2021_adjusted$mr <- 0

sapflow_usda_2021_adjusted[sapflow_usda_2021_adjusted$nodeId == "sapflow-usda-7",]$mr =  sapflow_usda_2021_adjusted[sapflow_usda_2021$nodeId == "sapflow-usda-7",]$meanHeatRatio + (1-as.numeric(usda_7_offset))

sapflow_usda_2021_adjusted[sapflow_usda_2021_adjusted$nodeId == "sapflow-usda-8",]$mr =  sapflow_usda_2021_adjusted[sapflow_usda_2021$nodeId == "sapflow-usda-8",]$meanHeatRatio + (1-as.numeric(usda_8_offset))

sapflow_usda_2021_adjusted[sapflow_usda_2021_adjusted$nodeId == "sapflow-usda-9",]$mr =  sapflow_usda_2021_adjusted[sapflow_usda_2021$nodeId == "sapflow-usda-9",]$meanHeatRatio + (1-as.numeric(usda_9_offset))

sapflow_usda_2021_adjusted[sapflow_usda_2021_adjusted$nodeId == "sapflow-usda-10",]$mr =  sapflow_usda_2021_adjusted[sapflow_usda_2021$nodeId == "sapflow-usda-10",]$meanHeatRatio + (1-as.numeric(usda_10_offset))



#SE 
sapflow_usda_2021_adjusted_se <- summarySE(sapflow_usda_2021_adjusted, measurevar="mr", groupvars=c("nodeId","hr"))




# TODO VPD SE


```

## Plot the adjusted with original 

```{r , echo=TRUE}

sapflow_usda_2021 %>% 
  mutate(node = as.factor(nodeId),hr = hour(dateP)) %>%
  mutate(mr = meanHeatRatio) %>%
  ggplot() + geom_point(aes(hr,mr,color=node) )  +
  facet_grid(~nodeId) +
  theme_classic() + labs(x="Date",y="Mean HR", title = "Node 7-10 USDA - Gravimetric Maize", caption="Source : dotmotelabs.com") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

```

```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate(node = as.factor(nodeId),hr = hour(dateP)) %>%
  ggplot() + geom_point(aes(hr,mr,color=node) )  +
  facet_grid(~nodeId) +
  theme_classic() + labs(x="Date",y="Mean HR", title = "Node 7-10 USDA - Gravimetric Maize", caption="Source : dotmotelabs.com") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

```

## Compare Original and Adjusted HR


```{r , echo=TRUE}

lm(sapflow_usda_2021_adjusted$meanHeatRatio~sapflow_usda_2021_adjusted$mr)

sapflow_usda_2021_adjusted %>% 
  ggplot() +
  geom_point(aes(meanHeatRatio,mr,color=nodeId)) +
  theme_minimal(base_size = 14)
```

## Heat pulse velocity

```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
#        filter(date(dateP) == "2020-12-01" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=sv,color=nodeId)) +
   geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = c(.17, .6))+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))

```
## Smooth

```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
  #filter(date(dateP) == "2021-03-20" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = hr, y=sv,color=nodeId)) +
  geom_smooth(aes(x = hr, y=sv,color=nodeId)) +
    theme_minimal() + 
  #scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  facet_wrap(~date(dateP) ) +

  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))
ggsave("figs/usda-graivimetric-Maize-vh-smoothed.png", width = 20, height = 20, units = "cm")

```

## Heat pulse velocity 1 to 15

```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
#        filter(date(dateP) == "2020-12-01" ) %>%
          filter(date(dateP) >= "2021-04-20" ) %>%
   filter(date(dateP) <= "2021-04-29" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=sv,color=nodeId)) +
   geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = c(.17, .6))+
    labs(color='Node ID',x="Time (Date)") +
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))
ggsave("figs/usda-graivimetric-Maize-stress-event.png", width = 20, height = 20, units = "cm")

```

## Mass flow


```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
   mutate (vs =1.7 * sv )%>% 
   mutate (flow = 3.14 * (1.5)^2 *vs )%>% 
   #     filter(date(dateP) == "2021-03-18" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=flow,color=nodeId)) +
   geom_line(aes(x = dateP, y=flow,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    #name = expression(~V[h] (cm~h^{"-1"}))
       # Features of the first axis
    name = expression("Sapflow rate"~J~(g~h^{"-1"})),
  ) +
  theme(legend.position = c(.17, .8))+
  labs(color='Node ID',x="Time (Date)") +
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))

```

## Get the scale data
```{r , echo=TRUE}


gravimetric_all <- readxl::read_xlsx('./data/Maize 4-16 to 4-30-21 Scale Data all MDT.xlsx')
gravimetric_all_raw <- readxl::read_xlsx('./data/Maize 4-16 to 4-30-21 Scale Data all MDT.xlsx',
                                         sheet ='Scales Raw Data' )
gravimetric_all_raw$time <- paste(lubridate::hour(gravimetric_all_raw$`Time (MDT)`),
                                      lubridate::minute(gravimetric_all_raw$`Time (MDT)`),
                                      lubridate::second(gravimetric_all_raw$`Time (MDT)`),sep = ":")
gravimetric_all_raw$datetime <- lubridate::as_datetime(trimws(paste(lubridate::as_date(gravimetric_all_raw$Date),
                                      gravimetric_all_raw$time," ")), tz="America/Denver")


```

## Map the patterns for Sunny days with Cloudy/Rainy Days - Use Agrimet. Use Sapflow velocity curves. Maybe above a threshold of 10 are viable days. 

```{r , echo=TRUE}

gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`)%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = datetime, y=weight,color=Dotmote)) +
   geom_line(aes(x = datetime, y=weight,color=Dotmote)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~Weight~(Kg))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")
ggsave("figs/gravi_plot.png", width = 20, height = 20, units = "cm")

```
```{r , echo=TRUE}

gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`)%>% 
     filter(date(datetime) >= "2021-04-28" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = datetime, y=weight,color=Dotmote)) +
   geom_line(aes(x = datetime, y=weight,color=Dotmote)) +
    theme_minimal(base_size = 18) + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~Weight~(Kg))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")
ggsave("figs/gravi_plot_28.png", width = 20, height = 20, units = "cm")

```

```{r , echo=TRUE}
sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
        filter(date(dateP) >= "2021-04-22" ) %>%
       filter(date(dateP) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=sv,color=nodeId)) +
   geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")
ggsave("figs/sap_plot.png", width = 20, height = 20, units = "cm")
```

```{r , echo=TRUE}
sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
        filter(date(dateP) >= "2021-04-22" ) %>%
       filter(date(dateP) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_smooth(aes(x = dateP, y=sv,color=nodeId)) +
   #geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal(base_size = 18) + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")
ggsave("figs/sap_smmoth_plot.png", width = 28, height = 15, units = "cm")
```

```{r , echo=TRUE}

usda_gravi <- gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`)%>% 
     filter(date(datetime) >= "2021-04-24" ) %>%
       filter(date(datetime) <= "2021-04-24" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = datetime, y=weight,color=Dotmote)) +
   geom_line(aes(x = datetime, y=weight,color=Dotmote)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~Weight~(Kg))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")
```

```{r , echo=TRUE}

sapflow_usda <- sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
       filter(date(dateP) == "2021-04-24" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=sv,color=nodeId)) +
   geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")

```

```{r , echo=TRUE}
library(patchwork)

usda_gravi / sapflow_usda
ggsave("figs/gravi_sapflow_usda.png", width = 20, height = 20, units = "cm")
```

# try the correlation
```{r , echo=TRUE}

usda_gravi_24 <- gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`, 
          hr = hour(datetime), mints = minute(datetime))%>% 
     filter(date(datetime) >= "2021-04-24" ) %>%
       filter(date(datetime) <= "2021-04-24" ) %>%
  as.data.frame()

usda_gravi_all <- gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`, 
          hr = hour(datetime), mints = minute(datetime))%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
  as.data.frame()

usda_gravi_24_8 <- gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`, 
          hr = hour(datetime), mints = minute(datetime))%>% 
     filter(date(datetime) >= "2021-04-24" ) %>%
       filter(date(datetime) <= "2021-04-24" ) %>%
  filter(Dotmote =='USDA8') %>%
  as.data.frame()

 #
sapflow_usda_24 <- sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
       filter(date(dateP) == "2021-04-24" ) %>%
  as.data.frame()

sapflow_usda_all <- sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
       filter(date(dateP) >= "2021-04-22" ) %>%
     filter(date(dateP) <= "2021-04-28" ) %>%
  as.data.frame()

sapflow_usda_24_8 <- sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
       filter(date(dateP) == "2021-04-24" & nodeId=='sapflow-usda-8') %>%
  as.data.frame()

```

```{r , echo=TRUE}

msv_s8 <- sapflow_usda_24_8 %>% 
   group_by(nodeId,hr) %>% 
 # arrange(nodeId, hr, sv) %>% 
  dplyr::summarise(m_sv = mean(sv)) %>%
  as.data.frame()  

msv_s_all <- sapflow_usda_all %>% 
   group_by(daten,nodeId,hr) %>% 
 # arrange(nodeId, hr, sv) %>% 
  dplyr::summarise(m_sv = mean(sv)) %>%
  as.data.frame()  

msv_s_all_7 <- sapflow_usda_all %>% 
  filter(nodeId=='sapflow-usda-7') %>%
   group_by(daten,nodeId,hr) %>% 
 # arrange(nodeId, hr, sv) %>% 
  dplyr::summarise(m_sv = mean(sv)) %>%
  as.data.frame()  

msv_s_all_8 <- sapflow_usda_all %>% 
  filter(nodeId=='sapflow-usda-8') %>%
   group_by(daten,nodeId,hr) %>% 
 # arrange(nodeId, hr, sv) %>% 
  dplyr::summarise(m_sv = mean(sv)) %>%
  as.data.frame()  

msv_s_all_9 <- sapflow_usda_all %>% 
  filter(nodeId=='sapflow-usda-9') %>%
   group_by(daten,nodeId,hr) %>% 
 # arrange(nodeId, hr, sv) %>% 
  dplyr::summarise(m_sv = mean(sv)) %>%
  as.data.frame()  

msv_s_all_10 <- sapflow_usda_all %>% 
  filter(nodeId=='sapflow-usda-10') %>%
   group_by(daten,nodeId,hr) %>% 
 # arrange(nodeId, hr, sv) %>% 
  dplyr::summarise(m_sv = mean(sv)) %>%
  as.data.frame()  


usda_gravi_24_8_weight_hrly <-  usda_gravi_24_8 %>%
  group_by(Dotmote,hr) %>%
    dplyr::summarise(m_weight= mean(weight)) %>%
as.data.frame()  

usda_gravi_all_weight_hrly <-  usda_gravi_all %>%
  group_by(daten,Dotmote,hr) %>%
    dplyr::summarise(m_weight= mean(weight)) %>%
as.data.frame()  

  
usda_gravi_24_8_weight_hrly_abs <- usda_gravi_24_8_weight_hrly %>%
  group_by(Dotmote,hr) %>%
  mutate(diff = abs(m_weight - lag(m_weight)) ) %>%
  as.data.frame()

usda_gravi_all_weight_hrly_abs <- usda_gravi_all_weight_hrly %>%
  group_by(daten,Dotmote,hr) %>%
  mutate(diff = abs(m_weight - lag(m_weight)) ) %>%
  order_by(daten,Dotmote,hr) %>%
  as.data.frame()

usda_gravi_all_weight_hrly_abs_7 <- usda_gravi_all_weight_hrly %>%
  filter(Dotmote == 'USDA7')  %>%
  group_by(daten,Dotmote,hr) %>%
  mutate(diff = abs(m_weight - lag(m_weight)) ) %>%
  as.data.frame()

usda_gravi_all_weight_hrly_abs_8 <- usda_gravi_all_weight_hrly %>%
  filter(Dotmote == 'USDA8')  %>%
  group_by(daten,Dotmote,hr) %>%
  mutate(diff = abs(m_weight - lag(m_weight)) ) %>%
  as.data.frame()

usda_gravi_all_weight_hrly_abs_9 <- usda_gravi_all_weight_hrly %>%
  filter(Dotmote == 'USDA9')  %>%
  group_by(daten,Dotmote,hr) %>%
  mutate(diff = abs(m_weight - lag(m_weight)) ) %>%
  as.data.frame()

usda_gravi_all_weight_hrly_abs_10 <- usda_gravi_all_weight_hrly %>%
  filter(Dotmote == 'USDA10')  %>%
  group_by(daten,Dotmote,hr) %>%
  mutate(diff = abs(m_weight - lag(m_weight)) ) %>%
  as.data.frame()


 summary( lm(usda_gravi_24_8_weight_hrly_abs$diff ~ msv_s8$m_sv))
```
# Plot hourly but one day
```{r , echo=FALSE}

ggplot() +
geom_point(data = usda_gravi_24_8_weight_hrly_abs, aes(x=hr,diff,color="GM")) +
geom_point(data= msv_s8, aes(x=hr, y=m_sv/100 , color="SV"))  +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
```

# Plot hourly but smooth it
```{r , echo=FALSE}

ggplot() +
geom_smooth(data = usda_gravi_24_8_weight_hrly_abs, aes(x=hr,diff,color="GM")) +
geom_smooth(data= msv_s8, aes(x=hr, y=m_sv/100 , color="SV"))  +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
```

# Compare , but one node one day

```{r , echo=FALSE}

ggplot() +
geom_point(data = usda_gravi_all_weight_hrly_abs_8[usda_gravi_all_weight_hrly_abs_8$daten=='2021-04-26',], aes(x=hr,diff,color=Dotmote)) +
geom_point(data= msv_s_all_8[msv_s_all_8$daten=='2021-04-26',], aes(x=hr, y=m_sv/100 , color=nodeId))  +
  facet_wrap(.~daten) +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
```
# All days, one node Node 7 perf

```{r , echo=FALSE}

coeff = 100
ggplot() +
  geom_point(data = usda_gravi_all_weight_hrly_abs_7[usda_gravi_all_weight_hrly_abs_7$daten > '2021-04-23',], aes(x=hr,diff,color=Dotmote)) +
  geom_point(data= msv_s_all_7[msv_s_all_7$daten > '2021-04-23', ], aes(x=hr, y=m_sv/coeff , color=nodeId))  +


  facet_wrap(.~daten) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Flow (kg/hr)" ,
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*(coeff), name=expression(~V[h] (cm~h^{"-1"})))
  ) +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
ggsave("figs/sap_gm_compare.png", width = 28, height = 15, units = "cm")
```


# All days, one node Node 8

```{r , echo=FALSE}

coeff = 100
ggplot() +
  geom_point(data = usda_gravi_all_weight_hrly_abs_8[usda_gravi_all_weight_hrly_abs_8$daten > '2021-04-23',], aes(x=hr,diff,color=Dotmote)) +
  geom_point(data= msv_s_all_8[msv_s_all_8$daten > '2021-04-23', ], aes(x=hr, y=m_sv/coeff , color=nodeId))  +


  facet_wrap(.~daten) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Flow (kg/hr)" ,
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*(coeff), name=expression(~V[h] (cm~h^{"-1"})))
  ) +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
```


# All days, one node Node 9

```{r , echo=FALSE}

coeff = 100
ggplot() +
  geom_point(data = usda_gravi_all_weight_hrly_abs_9[usda_gravi_all_weight_hrly_abs_9$daten > '2021-04-23',], aes(x=hr,diff,color=Dotmote)) +
  geom_point(data= msv_s_all_9[msv_s_all_9$daten > '2021-04-23', ], aes(x=hr, y=m_sv/coeff , color=nodeId))  +


  facet_wrap(.~daten) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Flow (kg/hr)" ,
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*(coeff), name=expression(~V[h] (cm~h^{"-1"})))
  ) +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
```

```{r , echo=FALSE}

coeff = 100
ggplot() +
  geom_point(data = usda_gravi_all_weight_hrly_abs_10[usda_gravi_all_weight_hrly_abs_10$daten > '2021-04-23',], aes(x=hr,diff,color=Dotmote)) +
  geom_point(data= msv_s_all_10[msv_s_all_10$daten > '2021-04-23', ], aes(x=hr, y=m_sv/coeff , color=nodeId))  +


  facet_wrap(.~daten) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Flow (kg/hr)" ,
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*(coeff), name=expression(~V[h] (cm~h^{"-1"})))
  ) +
  theme_minimal(base_size = 18) +
  labs(color="",x=" Time (Hour)" , y="")
```

```{r , echo=FALSE}
# Adjust the Sapflow

coeff <- 1

msv_s8 %>% 
  ggplot() +
   geom_point(aes(x = hr, y=m_sv,color=nodeId)) +
   geom_line(aes(x = hr, y=m_sv,color=nodeId)) +
   geom_point( data=usda_gravi_24_8_weight_hrly,aes(x=hr,y=m_weight / coeff,color='Weigh')) +
  #geom_smooth(aes(dateP,sv),span = 0.1, se = FALSE) +
  #geom_smooth(aes(x = dateP, y=meanhr,color=as.factor(nodeID)),method = loess, span="0.0008",se = TRUE)  +
    theme_minimal() + 
  #scale_x_datetime(date_minor_breaks = "1 day") + 
  # labs(subtitle="",title=expression(~italic(Corn~~USDA)), y =expression(~V[h] (cm~h^{"-1"})),x="Date",
  #              colour = "Node ID") +
  # Divide by 10 to get the same range than the temperature
  
  scale_y_continuous(
    
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"})),
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Flow (g/s)")
  ) +
  theme(legend.position = c(.17, 1))+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))
```



```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP))%>% 
       #filter(date(dateP) == "2021-04-24" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = dateP, y=sv,color=nodeId)) +
   geom_line(aes(x = dateP, y=sv,color=nodeId)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position = c(.17, .6))+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1))

```



```{r , echo=TRUE}

 gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`,day= day(datetime),hour=hour(datetime))%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = hour, y=weight,color=Dotmote)) +
   geom_line(aes(x = hour, y=weight,color=Dotmote)) +
    theme_minimal() + 
  #scale_x_datetime(date_minor_breaks = "1 day") + 
  facet_wrap(~day) +

  scale_y_continuous(
    # Features of the first axis
    name = expression(~Weight~(Kg))

  ) +
  theme(legend.position="bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x=" Time (Hour)")

 gravitim <- gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`,day= day(datetime),hour=hour(datetime))%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = hour, y=weight,color=Dotmote)) +
   geom_line(aes(x = hour, y=weight,color=Dotmote)) +
    theme_minimal() + 
  #scale_x_datetime(date_minor_breaks = "1 day") + 
  facet_wrap(~day) +

  scale_y_continuous(
    # Features of the first axis
    name = expression(~Weight~(Kg))

  ) +
  theme(legend.position="bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x=" Time (Hour)")

```



```{r , echo=TRUE}

sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP), 
          day=day(dateP),hour=hour(dateP))%>% 
       #filter(date(dateP) == "2021-04-24" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
       filter(date(dateP)  >= "2021-04-22" ) %>%
       filter(date(dateP)  <= "2021-04-28" ) %>%
  ggplot() +
   geom_point(aes(x = hour, y=sv,color=nodeId)) +
   geom_line(aes(x = hour, y=sv,color=nodeId)) +
    theme_minimal() + 
  #scale_x_datetime(date_minor_breaks = "1 day") + 
facet_wrap(~day)+
  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position="bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Hour)")

saptim <-  sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP), 
          day=day(dateP),hour=hour(dateP))%>% 
       #filter(date(dateP) == "2021-04-24" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
       filter(date(dateP)  >= "2021-04-22" ) %>%
       filter(date(dateP)  <= "2021-04-28" ) %>%
  ggplot() +
   geom_point(aes(x = hour, y=sv,color=nodeId)) +
   geom_line(aes(x = hour, y=sv,color=nodeId)) +
    theme_minimal() + 
  #scale_x_datetime(date_minor_breaks = "1 day") + 
facet_wrap(~day)+
  scale_y_continuous(
    # Features of the first axis
    name = expression(~V[h] (cm~h^{"-1"}))

  ) +
  theme(legend.position="bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Hour)")

```

```{r , echo=TRUE}
library(patchwork)

gravitim / saptim
ggsave("figs/gravi_sapflow_timeseries_usda.png", width = 20, height = 20, units = "cm")
```


# Explore relationship between Sapflow and Gravimetric 

```{r , echo=TRUE}

blockDataSapflow <- sapflow_usda_2021_adjusted %>% 
  mutate (sv = log(mr) *3600 * (0.0025/0.6) ,daten = date(dateP), 
          day=day(dateP),hour=hour(dateP))%>% 
       #filter(date(dateP) == "2021-04-24" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
       filter(date(dateP)  >= "2021-04-22" ) %>%
       filter(date(dateP)  <= "2021-04-28" ) %>% 
  as.data.frame()
blockDataGavimetric<- gravimetric_all_raw %>% 
  mutate (daten = date(datetime),weight = `Weight(kg)`,day= day(datetime),hour=hour(datetime))%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
   as.data.frame()

# Take values from 6 to 18 for all the plants
filteredSapflow <- blockDataSapflow %>% 
  #filter(hour %in% c(6:18)) %>% 
  dplyr::group_by(day,hour) %>%
  dplyr::summarize(mhr=mean(mr)) %>%
  select(day,hour,mhr) %>%
  as.data.frame()
filteredGravimetric <- blockDataGavimetric %>% 
  #filter(hour %in% c(6:18)) %>% 
    dplyr::group_by(day,hour) %>%
  dplyr::summarize(mw=mean(weight)) %>%
  select(day,hour,mw) %>%
  as.data.frame()
```


```{r , echo=TRUE}

combinedFiltered <-  filteredSapflow
combinedFiltered$mw <- filteredGravimetric[c(1:167),]$mw

# Take values from 6 to 18 for all the plants
ggplot() +
  geom_point(data=combinedFiltered,aes(mw,mhr,color=as.factor(hour))) +
  theme_minimal(base_size = 14)+ 
  labs(x="Weight (kg)", y="Vh") +
  labs(color="Hour")
ggsave("figs/gravi_sapflow_compare.png", width = 20, height = 20, units = "cm")
```


```{r , echo=TRUE}
summary(lm(data = combinedFiltered, mhr~mw))
```

# check the temp

```{r , echo=TRUE}
climate_gravimetric_all <- readxl::read_xlsx('./data/Greenhouse 5 min data 4-16 to 4-30-21 Maize.xlsx')
head(climate_gravimetric_all)
climate_gravimetric_all$time <- paste(lubridate::hour(climate_gravimetric_all$Time),
                                      lubridate::minute(climate_gravimetric_all$Time),
                                      lubridate::second(climate_gravimetric_all$Time),sep = ":")
climate_gravimetric_all$datetime <- lubridate::as_datetime(trimws(paste(lubridate::as_date(climate_gravimetric_all$Date),
                                      climate_gravimetric_all$time," ")), tz="America/Denver")

```

# plot for the day

```{r , echo=TRUE}

climate_gravimetric_all %>% 
  mutate (daten = date(datetime),tempF = `Climate Temperature(°F)`)%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
 # filter(nodeId=="sapflow-dotmote-labs-2" )  %>% 
  ggplot() +
   geom_point(aes(x = datetime, y=tempF)) +
   geom_line(aes(x = datetime, y=tempF)) +
    theme_minimal() + scale_x_datetime(date_minor_breaks = "1 day") + 

  scale_y_continuous(
    # Features of the first axis
    name = expression(~Temperature~(F))

  ) +
  theme(legend.position = "bottom")+
guides(fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype=guide_legend(keywidth = 3, keyheight = 1)) +
  labs(color="Node",x="Time (Date)")
ggsave("figs/gravi_climate.png", width = 20, height = 20, units = "cm")

```


VPD
```{r , echo=FALSE}

#(0.611 * 10^ (7.5 * T / (237 +T)) * (1− RH/100)
 
climate_gravimetric_all %>% 
  mutate (daten = date(datetime),tempF = `Climate Temperature(°F)`,rh = `Climate Humidity(%Rh)`)%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
  mutate(vpd = (0.611 * 10^((7.5 * tempF )/(237 + tempF)) * ( 1- (rh/100)))) %>% 
  ggplot( aes(x=datetime, y=vpd))  +
    geom_line(color='red') +
    geom_point() +
    labs(x="Time (Date)", y="VPD (kPa)",colour = "Hibiscus",title = "") +
   theme_minimal(base_size = 18) +
  theme(legend.position = c(1, 0.0050))

ggsave("figs/gravi_climate_vpd.png", width = 20, height = 20, units = "cm")
```

```{r , echo=FALSE}

#(0.611 * 10^ (7.5 * T / (237 +T)) * (1− RH/100)
 
climate_gravimetric_all %>% 
  mutate (daten = date(datetime),tempF = `Climate Temperature(°F)`,
          rh = `Climate Humidity(%Rh)`,
          sr= `TCAP; Broadcast Information; Outdoor Light Energy(W/m²)`)%>% 
     filter(date(datetime) >= "2021-04-22" ) %>%
       filter(date(datetime) <= "2021-04-28" ) %>%
  mutate(vpd = (0.611 * 10^((7.5 * tempF )/(237 + tempF)) * ( 1- (rh/100)))) %>% 
  ggplot( aes(x=datetime, y=sr))  +
    geom_line(color='red') +
    geom_point() +
    labs(x="Time (Date)", y="Outdoor Light Energy(W/m²)",colour = "Hibiscus",title = "") +
   theme_minimal(base_size = 18) +
  theme(legend.position = c(1, 0.0050))

ggsave("figs/gravi_climate_energy.png", width = 20, height = 20, units = "cm")
```
